Probabilistic bioregion models suggest seabird assemblages are linked to the major boundary current off eastern Australia

### 2023-06-27 docs/ms-draft.Rmd
fixed tables 1 (cfs) and 2 (-Inf)

---
title: "Probabilistic bioregion models suggest seabird assemblages are linked to the major boundary current off eastern Australia"
# "Probabilistic bioregion models suggest seabird assemblages are linked to large scale oceanography off eastern Australia"
# author: Nicholas W Daudt
# date: '`r format(Sys.Date())`'
author:
  - Nicholas W. Daudt:
      email: nicholaswdaudt@gmail.com
      institute: [MarSciUO]
      correspondence: true
  - Eric J. Woehler:
      institute: [ASG]
      correspondence: false
  - Matthew R. Schofield:
      institute: [MathStatUO]
      correspondence: false
  - Robert O. Smith:
      institute: [MarSciUO]
      correspondence: false
  - Leandro Bugoni:
      institute: [FURG]
      correspondence: false
  - William J. Rayment:
      institute: [MarSciUO]
      correspondence: false
institute:
  - MarSciUO: Department of Marine Science, University of Otago, Aotearoa New Zealand
  - ASG: Australasian Seabird Group, BirdLife Australia, Hobart, Australia
  - MathStatUO: Department of Mathematics and Statistics, University of Otago, Aotearoa New Zealand
  - FURG: Institute of Biological Sciences, Universidade Federal do Rio Grande--FURG, Rio Grande, RS, Brazil
abstract: |
  **Aim**: To identify assemblages of seabird species at large scale, probabilistically, and related them to environmental conditions off eastern Australia, a world seabird hotspot.
  **Methods**: We used seabird at-sea ship-based survey data collected over 37° latitude and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and counts). RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within it but different from other regions. These groups also vary according to covariates, on which we included oceanographic static and climatic, as well as physiographic layers.
  **Results**: All models suggested two large scale assemblages ('northern' and 'southern' assemblages), except for the abundance-autumn model which indicated three groups. Model results seem to agree in a biogeographic barrier at ~32°S, where the East Australian Current mostly dissipates to the east and loses force. Furthermore, sea surface temperature and, in one case, sea surface salinity, were selected in all final models, further indicating close relationship between seabird assemblages and water masses. Results from both data types, presence-absence and counts resulted in similar spatial and species profile patterns.
  **Conclusion**: Two seabird assemblages were clearly identified from RCP models, suggesting the persistence of these groups at temporal and spatial scales. Given the rapid change that the East Australian Current is experiencing, and its importance on driving seabirds' distributions, the methods applied in our study could be replicated in future analyses to assess possible spatial/temporal changes in seabird assemblages, including on its structure.
output: 
  # html_document: default
  # word_document:
  #   reference_docx: "../docs/docx-template.docx"
  # officedown::rdocx_document:
  #   reference_docx: "../docs/docx-template.docx"
  # rticles::sim_article:
  #   keep_tex: TRUE
  bookdown::word_document2:
    pandoc_args:
      - --lua-filter=../docs/scholarly-metadata.lua
      - --lua-filter=../docs/author-info-blocks.lua
      - --reference-doc=../docs/docx-template.docx 
bibliography: ../references.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
---

@@

```{r setup, include = FALSE, echo = FALSE}
library(gt)
# library(officedown)
# library(officer)
```

@@

Seabird assemblages have been described and related to environmental characteristics worldwide. Considering the hierarchical nature of oceanographic processes and its resulting features (**Haury78**), it is not surprising that the distribution of high mobile seabirds are linked differently across scales [@hunt1987]. In the marine environment, large scale oceanographic features (water masses and currents) act as 'biomes' and influence seabird assemblage structure--i.e. the occurrence of species and its relative numerical composition [@ribic1997; @hyrenbach2007; @commins2014]. On the other hand, at smaller scales, local abundance is related to temporary processes that aggregate prey and therefore predators, such as fronts and eddies [@weimerskirch2007; @scales2014]. Such smaller-scale processes, although ephemeral, are primarily driven by physical and atmospheric forces and are thus somewhat predictable in space and time, providing consistent feeding areas for marine predators when they occur [@davoren2013; @hazen2013]. Noticeably, sea surface temperature is a key driver of distribution of many seabird species, at individual- and assemblage-levels, as well as at large (water masses) and smaller (fronts) scales [@dunlop1988; @ribic1997; @quillfeldt2015]. Nonetheless, many regions still lack information regarding multiple seabird species distribution and how they assemble in space and time.

@@

```{r fig1-study-area, fig.cap = "**FIGURE 1.** Study area showing the main ocean currents (a), and seabird records by season on top of the 1° latitude × 1° longitude grid cells (b). In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue color. The EAC system originates from the South Equatorial Current (SEC) and its branches, North Caledonian Jet (NCJ) and South Caledonian Jet (SCJ), drawn in red. The EAC 'eastern extension' is an eddy field (blue shaded area) often mentioned as the 'Tasman Front', from which the East Auckland Current originates (EAUC, in green). Figure (a) is adapted after Oke, Roughan, et al. (2019); for colour references see the online version.", fig.align = "center", echo = FALSE}

knitr::include_graphics("../results/Fig1_study-area-w-oceanogr.png")
```

@@

data.frame(
  Variable = c("Sea Surface Temperature", "Sea Surface Temperature gradient", "Sea Surface Salinity", "Mixed Layer Depth", "8-day average Chlorophyll-a", "Log10 8-day average Chlorophyll-a", "Eddy Kinetic Energy", "Climatic mean EKE", "Climatic standard deviation EKE", "Climatic mean SSTgrad", "Bathymetry", "Slope", "Distance from coast"),
  Abbreviation = c("sst", "sstgrad", "sss", "mld", "chl8", "log10_chl8", "eke", "clim_eke", "clim_ekesd", "clim_sstgrad", "bat", "slope", "dist_coast"),
  Units = c("°Celsius", "delta-°Celsius", "PSU", "m", "mg/m^3", "log10(chl8)", "m/s", "mean(m/s)/season/10 years", "sd(m/s)/season/10 years", "mean(delta-°Celsius)/season/10 years", "m", "degrees", "km"),
  `Spatial resolution` = c("0.25° × 0.25°", "0.25° × 0.25°", "0.083° × 0.083°", "0.083° × 0.083°", "0.036° × 0.036°", "0.036° × 0.036°", "cf", "cf", "cf", "0.25° × 0.25°", "cf", "cf", "NA"),
  `Temporal resolution` = c("daily", "daily", "daily", "daily", "8-days composite", "8-days composite", "daily", "seasonal climatology", "seasonal climatology", "seasonal climatology", "NA", "NA", "NA"),
  Source = c("Huang et al. (2021)", "Derived from Huang et al. (2021)", "E.U. Copernicus Marine Service Information", "E.U. Copernicus Marine Service Information", "ERDDAP/NOAA", "Derived from ERDDAP/NOAA", "IMOS/AODN", "Derived from IMOS/AODN", "Derived from IMOS/AODN", "Derived from Huang et al. (2021)", "GEBCO", "Derived from GEBCO", "Calculated in R (see main text)")
) |>
  gt::gt() |>
  ## SST & SSTgrad
  gt::tab_footnote(footnote = 
                     gt::md("[NOAA OISST](https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html)"),
                   locations = gt::cells_body(columns = Source, rows = c(1:2, 10))) |>
  ## SSS & MLD
  gt::tab_footnote(footnote = 
                     gt::md("2011--2019: [GLOBAL_MULTIYEAR_PHY_001_030](https://doi.org/10.48670/moi-00021); 2020--2021: [GLOBAL_ANALYSIS_FORECAST_PHY_001_024](https://doi.org/10.48670/moi-00016)"),
                   locations = gt::cells_body(columns = Source, rows = 3:4)) |>
  ## CHL8 & log10-CHL8
  gt::tab_footnote(footnote = 
                     gt::md("[ERDDAP/NOAA](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html)"),
                   locations = gt::cells_body(columns = Source, rows = 5:6)) |>
  ## EKE and derivations
  gt::tab_footnote(footnote = 
                     gt::md("[IMOS/AODN](https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set"),
                   locations = gt::cells_body(columns = Source, rows = 7:9)) |>
  ## GEBCO & slope
  gt::tab_footnote(footnote = 
                     gt::md("[GEBCO](https://www.gebco.net/data_and_products/gridded_bathymetry_data/)"),
                   locations = gt::cells_body(columns = Source, rows = 11:12)) |>
  ## Improve cols names
  gt::cols_label(
    Variable = gt::md("**Variable**"),
    Abbreviation = gt::md("**Abbreviation**"),
    Units = gt::md("**Units**"),
    `Spatial.resolution` = gt::md("**Spatial resolution**"), 
    `Temporal.resolution` = gt::md("**Temporal resolution**"),
    Source = gt::md("**Source**")) |>
  ## Table title
  gt::tab_header(title = gt::md("**TABLE 1.** Environmental data used in the Region of Common Profile models applied to seabird data off eastern Australia, their spatial/temporal resolutions, units, and sources"))

@@

<!-- summer \~ bat+slope+eke+sst+sss+mld+sst_grad -->

<!-- autumn \~ bat+slope+eke+chl+sst+sss+mld+sst_grad+clim_sst_grad -->

<!-- winter \~ bat+slope+eke+sst+sss+mld+sst_grad+clim_sst_grad -->

<!-- spring \~ bat+slope+dist_coast+eke+chl+sst+sss+mld+sst_grad+clim_sst_grad+clim_eke_mean -->

@@

A total of 10,261 seabird records were obtained. They summed to 142,646 individuals from 80 species, belonging to 11 families and five orders. The five most abundant species summed 83% of total birds counted (43.0% *Ardenna tenuirostris*, 17.2% *Pachyptila turtur*, 14.3% *Ardenna pacifica*, 5.2% *Onychoprion fuscatus*, 3.2% *Thalassarche cauta*), whereas the five most frequent species were *Ardenna pacifica* (frequency of occurrence = 35.3%), *Thalassarche cauta* (33.2%), *Sula leucogaster* (25.4%), *Sula sula* (25.4%) and *Thalassarche melanophris* (25%) (**Fig. S1 - FO/nOCC/FN**). An increase in species richness towards the south was clear in all but the winter season (**Fig S2. - spp richness/total abund**). Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' (**Fig. S2**). However, one cell in the Coral Sea summed to numbers of the same order of magnitude as off Tasmania in summer.

@@

All models suggested two RCPs as the optimal grouping number, except for P/A-autumn, which indicated three RCPs as the best fit (**Fig S3. - multifit plots**). The best fitting models generally had **1--3** covariates selected, although the abundance-autumn model had five covariates (**Table 2 - full-model/selected model**). All the best models selected SST or SSS, variables essential for characterising water masses. The best models show a good fit with no apparent deviances in assumptions (**Fig. S4 - residuals**).

@@

```{r fig2-point-pred, fig.cap = "**FIGURE 2.** Spatialized point predictions (hard-classes) for seabird Region of Common Profile (RCP) models, by season, based on presence-absence (a) and abundance (b) models off eastern Australia. Each grid was assigned to the RCP group with highest membership probability.", fig.align = "center", echo = FALSE}

@@

```{r fig3-prob-maps, fig.cap = "**FIGURE 3.** Predicted probabilities of Region of Common Profiles membership for each grid, based on selected covariates (see Table 2) of spring presence-absence (a) and abundance (b) models for seabird assemblages off eastern Australia. Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}

@@

@@

```{r fig5-spp-profiles, fig.cap = "**FIGURE 5.** Species profiles of seabird assembalges off eastern Australia given by spring presence-absence (a) and abundance (b) Region of Common Profile models. Note that values in (b) are log10 transformed. Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}

@@

Although the results suggest consistent patterns, we should point out some caveats of our analyses. Seabird data were gathered opportunistically while taking part on several projects, whose sampling design were not planned to account for seabird spatial ecology specifically. As a result, there are spatial discontinuities on the dataset, which might have limited the ability of our models to detect the relationships between assemblages and environmental predictors more robustly. Even though Australia is a hotspot for seabirds [@karpouzi2007; @ramírez2017], there is not much information regarding seabird distribution to date [@mott2018; @bernard2021]. Dedicated seabird effort could thus benefit future studies advancing seabird research in this region, as well as improving patterns described here. By increasing the amount of data, future studies could set the grid cells at higher resolution, and potentially reveal assemblages at a finer scale. In addition, ideally, we could have had more information within each grid cell by season (our sampling units). Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species on most of them (**Fig S8 - rarefaction curves**).

>>>>

---
title: "Probabilistic bioregion models suggest seabird assemblages are linked to the major boundary current off eastern Australia"
author:
  - Nicholas W. Daudt:
      email: nicholaswdaudt@gmail.com
      institute: [MarSciUO]
      correspondence: true
  - Eric J. Woehler:
      institute: [ASG]
      correspondence: false
  - Matthew R. Schofield:
      institute: [MathStatUO]
      correspondence: false
  - Robert O. Smith:
      institute: [MarSciUO]
      correspondence: false
  - Leandro Bugoni:
      institute: [FURG]
      correspondence: false
  - William J. Rayment:
      institute: [MarSciUO]
      correspondence: false
institute:
  - MarSciUO: Department of Marine Science, University of Otago, Aotearoa New Zealand
  - ASG: Australasian Seabird Group, BirdLife Australia, Hobart, Australia
  - MathStatUO: Department of Mathematics and Statistics, University of Otago, Aotearoa New Zealand
  - FURG: Institute of Biological Sciences, Universidade Federal do Rio Grande--FURG, Rio Grande, RS, Brazil
abstract: |
  **Aim**: To identify assemblages of seabird species at large scale, probabilistically, and related them to environmental conditions off eastern Australia, a world seabird hotspot.
  **Methods**: We used seabird at-sea ship-based survey data collected over 37° latitude and fitted Region of Common Profile (RCP) mixture models for each season and type of data (presence-absence and counts). RCP groups are defined as a set of regions where the probability of encountering a specific species profile is constant within it but different from other regions. These groups also vary according to covariates, on which we included oceanographic static and climatic, as well as physiographic layers.
  **Results**: All models suggested two large scale assemblages ('northern' and 'southern' assemblages), except for the abundance-autumn model which indicated three groups. Model results seem to agree in a biogeographic barrier at ~32°S, where the East Australian Current mostly dissipates to the east and loses force. Furthermore, sea surface temperature and, in one case, sea surface salinity, were selected in all final models, further indicating close relationship between seabird assemblages and water masses. Results from both data types, presence-absence and counts resulted in similar spatial and species profile patterns.
  **Conclusion**: Two seabird assemblages were clearly identified from RCP models, suggesting the persistence of these groups at temporal and spatial scales. Given the rapid change that the East Australian Current is experiencing, and its importance on driving seabirds' distributions, the methods applied in our study could be replicated in future analyses to assess possible spatial/temporal changes in seabird assemblages, including on its structure.
output: 
  # rticles::sim_article:
  #   keep_tex: TRUE
  bookdown::word_document2:
    pandoc_args:
      - --lua-filter=../docs/scholarly-metadata.lua
      - --lua-filter=../docs/author-info-blocks.lua
      - --reference-doc=../docs/docx-template.docx 
bibliography: ../references.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
---

@@

```{r setup, include = FALSE, echo = FALSE}
library(gt)
```

@@

Seabird assemblages have been described and related to environmental characteristics worldwide. Considering the hierarchical nature of oceanographic processes and its resulting features [@haury1978patterns], it is not surprising that the distribution of high mobile seabirds are linked differently across scales [@hunt1987]. In the marine environment, large scale oceanographic features (water masses and currents) act as 'biomes' and influence seabird assemblage structure--i.e. the occurrence of species and its relative numerical composition [@ribic1997; @hyrenbach2007; @commins2014]. On the other hand, at smaller scales, local abundance is related to temporary processes that aggregate prey and therefore predators, such as fronts and eddies [@weimerskirch2007; @scales2014]. Such smaller-scale processes, although ephemeral, are primarily driven by physical and atmospheric forces and are thus somewhat predictable in space and time, providing consistent feeding areas for marine predators when they occur [@davoren2013; @hazen2013]. Noticeably, sea surface temperature is a key driver of distribution of many seabird species, at individual- and assemblage-levels, as well as at large (water masses) and smaller (fronts) scales [@dunlop1988; @ribic1997; @quillfeldt2015]. Nonetheless, many regions still lack information regarding multiple seabird species distribution and how they assemble in space and time.

@@

```{r fig1-study-area, fig.cap = "**FIGURE 1.** Study area showing the main ocean currents (a), and seabird records made by season on top of the 1° latitude × 1° longitude grid cells (b). In (a), the East Australian Current (EAC) system is highlighted with its acronyms in bold and drawn in blue color. The EAC system originates from the South Equatorial Current (SEC) and its branches, North Caledonian Jet (NCJ) and South Caledonian Jet (SCJ), drawn in red. The EAC 'eastern extension' is an eddy field (blue shaded area) often mentioned as the 'Tasman Front', from which the East Auckland Current originates (EAUC, in green). Figure (a) is adapted after Oke, Roughan, et al. (2019); for colour references see the online version.", fig.align = "center", echo = FALSE}

# knitr::include_graphics("../results/Fig1_study-area-w-oceanogr.png")

knitr::include_graphics("../EDA/Fig1-study-area-raw-counts.png")
```

@@

data.frame(
  Variable = c("Sea Surface Temperature", "Sea Surface Temperature gradient", "Sea Surface Salinity", "Mixed Layer Depth", "8-day average Chlorophyll-a", "Log10 8-day average Chlorophyll-a", "Eddy Kinetic Energy", "Climatic mean EKE", "Climatic standard deviation EKE", "Climatic mean SSTgrad", "Bathymetry", "Slope", "Distance from coast"),
  Abbreviation = c("sst", "sstgrad", "sss", "mld", "chl8", "log10_chl8", "eke", "clim_eke", "clim_ekesd", "clim_sstgrad", "bat", "slope", "dist_coast"),
  Units = c("°Celsius", "delta-°Celsius", "PSU", "m", "mg/m^3", "log10(chl8)", "m/s", "mean(m/s)/season/10 years", "sd(m/s)/season/10 years", "mean(delta-°Celsius)/season/10 years", "m", "degrees", "km"),
  `Spatial resolution` = c("0.25° × 0.25°", "0.25° × 0.25°", "0.083° × 0.083°", "0.083° × 0.083°", "0.036° × 0.036°", "0.036° × 0.036°", "0.2° × 0.2°", "0.2° × 0.2°", "0.2° × 0.2°", "0.25° × 0.25°", "0.004° × 0.004°", "0.004° × 0.004°", "NA"),
  `Temporal resolution` = c("daily", "daily", "daily", "daily", "8-days composite", "8-days composite", "daily", "seasonal climatology", "seasonal climatology", "seasonal climatology", "NA", "NA", "NA"),
  Source = c("Huang et al. (2021)", "Derived from Huang et al. (2021)", "E.U. Copernicus Marine Service Information", "E.U. Copernicus Marine Service Information", "ERDDAP/NOAA", "Derived from ERDDAP/NOAA", "IMOS/AODN", "Derived from IMOS/AODN", "Derived from IMOS/AODN", "Derived from Huang et al. (2021)", "GEBCO", "Derived from GEBCO", "Calculated in R (see main text)")
) |>
  gt::gt() |>
  ## SST & SSTgrad
  gt::tab_footnote(footnote = 
                     gt::md("[NOAA OISST](https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html)"),
                   locations = gt::cells_body(columns = Source, rows = c(1:2, 10))) |>
  ## SSS & MLD
  gt::tab_footnote(footnote = 
                     gt::md("2011--2019: [GLOBAL_MULTIYEAR_PHY_001_030](https://doi.org/10.48670/moi-00021); 2020--2021: [GLOBAL_ANALYSIS_FORECAST_PHY_001_024](https://doi.org/10.48670/moi-00016)"),
                   locations = gt::cells_body(columns = Source, rows = 3:4)) |>
  ## CHL8 & log10-CHL8
  gt::tab_footnote(footnote = 
                     gt::md("[ERDDAP/NOAA](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html)"),
                   locations = gt::cells_body(columns = Source, rows = 5:6)) |>
  ## EKE and derivations
  gt::tab_footnote(footnote = 
                     gt::md("[IMOS/AODN](https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set"),
                   locations = gt::cells_body(columns = Source, rows = 7:9)) |>
  ## GEBCO & slope
  gt::tab_footnote(footnote = 
                     gt::md("[GEBCO](https://www.gebco.net/data_and_products/gridded_bathymetry_data/)"),
                   locations = gt::cells_body(columns = Source, rows = 11:12)) |>
  ## Improve cols names
  gt::cols_label(
    Variable = gt::md("**Variable**"),
    Abbreviation = gt::md("**Abbreviation**"),
    Units = gt::md("**Units**"),
    `Spatial.resolution` = gt::md("**Spatial resolution**"), 
    `Temporal.resolution` = gt::md("**Temporal resolution**"),
    Source = gt::md("**Source**")) |>
  ## Table title
  gt::tab_header(title = gt::md("**TABLE 1.** Environmental data used in the Region of Common Profile models applied to seabird data off eastern Australia, their spatial/temporal resolutions, units, and sources"))

@@

@@

A total of 10,261 seabird records were obtained. They summed to 142,646 individuals from 80 species, belonging to 11 families and five orders. The five most abundant species summed 83% of total birds counted (43.0% *Ardenna tenuirostris*, 17.2% *Pachyptila turtur*, 14.3% *Ardenna pacifica*, 5.2% *Onychoprion fuscatus*, 3.2% *Thalassarche cauta*), whereas the five most frequent species were *Ardenna pacifica* (frequency of occurrence = 35.3%), *Thalassarche cauta* (33.2%), *Sula leucogaster* (25.4%), *Sula sula* (25.4%) and *Thalassarche melanophris* (25%) (**Fig. S1**). An increase in species richness towards the south was clear in all but the winter season (**Fig S2**). Noticeably higher counts were made around Tasmania, possibly due to the 'colony effect' (**Fig. S2**). However, one cell in the Coral Sea summed to numbers of the same order of magnitude as off Tasmania in summer.

@@

All models suggested two RCPs as the optimal grouping number, except for P/A-autumn, which indicated three RCPs as the best fit (**Fig S3**). The best fitting models generally had 1--2 covariates selected, although the abundance-autumn model had five covariates (**Table 2**). All the best models selected SST or SSS, variables essential for characterising water masses. The best models show a good fit with no apparent deviances in assumptions (**Fig. S4**).

```{r table2, echo = FALSE, message = FALSE}

df <- 
  rbind(
    list.files(path = "../results/Bernoulli", pattern = "*04_full_and_best_models.csv", full.names = TRUE),
    list.files(path = "../results/NegBin", pattern = "*04_full_and_best_models.csv", full.names = TRUE)) %>% 
  purrr::map_df(~readr::read_csv(.)) %>% 
  dplyr::select(- `...1`) %>% 
  dplyr::mutate(data_type = factor(data_type, levels = c("bernoulli", "neg_bin")),
                season = factor(season, levels = c("summer", "autumn", "winter", "spring")),
                model = factor((ifelse(model == "full", "a", "b")), levels = c("a", "b")),
                bic = round(bic, digits = 2)) %>%
  dplyr::group_by(data_type) %>% 
  dplyr::arrange(data_type, season) %>% 
  as.data.frame(.) %>% 
  dplyr::ungroup(.)

# Correct a value in a cell
# Full abundance-autumn model did not fit properly, returning a value of -Inf
df[df$data_type=='neg_bin' & df$season=='autumn' & df$model=='a',]$bic <- "-Inf (n.f.)"

# A few tweaks to make the table look better
df$data_type <- c("Bernoulli", rep("", 7), "Negative Binomial", rep("", 7))
df$season <- rep(c("Summer", "", "Autumn", "", "Winter", "", "Spring", ""), 2)

gt::gt(df)|>
  ## Improve cols names
  gt::cols_label(
    data_type = gt::md("**Data type**"),
    season = gt::md("**Season**"),
    model = gt::md(" "),          ## To be excluded in the Word doc
    model_specification = gt::md("**Model specification**"), 
    bic = gt::md("**BIC**")) |>
  ## Table title
  gt::tab_header(title = gt::md("**TABLE 2.** Full (a) and best (b) Region of Common Profile Models applied to seabird data off eastern Australia, for both data types used (presence-absence, based on a Bernoulli error distribution; abundance, based on a Negative Binomial error distribution)"))

rm("df")
## Can't figure out why this code below doesn't work...
# kableExtra::kbl(df, booktabs = T, align = "l", linesep = "") %>% 
#   kableExtra::column_spec(1, bold=T) %>%
#   kableExtra::collapse_rows(columns = c(1:3), 
#                             latex_hline = 'custom', 
#                             custom_latex_hline = 1:3, 
#                             row_group_label_position = 'stack')
```

@@

```{r fig2-point-pred, fig.cap = "**FIGURE 2.** Spatialized point predictions (hard-classes) for seabird assemblages (Region of Common Profile; RCP), by season, based on presence-absence (a) and abundance (b) models, off eastern Australia. Each grid was assigned to the RCP group with highest membership probability.", fig.align = "center", echo = FALSE}

@@

```{r fig3-prob-maps, fig.cap = "**FIGURE 3.** Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from spring presence-absence (a) and abundance (b) models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides. Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}

@@

```{r fig4-partial-plots, fig.cap = "**FIGURE 4.**  Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}

knitr::include_graphics("../results/Fig4_partial-plots-spring.png")
```

@@

```{r fig5-spp-profiles, fig.cap = "**FIGURE 5.** Species profiles of seabird assembalges (Region of Common Profile; RCP) off eastern Australia for spring presence-absence (a) and abundance (b) models. Values are the average and confidence intervals of probability of occurrence (a) and predicted mean abundance (b) for each species, based on 1000 Bayesian bootstraps. Values in (b) were log10-transformed to accomodate the high variation between species. Results for other seasons and models can be found in the Supplementary Material.", fig.align = "center", echo = FALSE}

@@

Although the results suggest consistent patterns, we should point out some caveats of our analyses. Seabird data were gathered opportunistically while taking part on several projects, whose sampling design were not planned to account for seabird spatial ecology specifically. As a result, there are spatial discontinuities on the dataset, which might have limited the ability of our models to detect the relationships between assemblages and environmental predictors more robustly. Even though Australia is a hotspot for seabirds [@karpouzi2007; @ramírez2017], there is not much information regarding seabird distribution to date [@mott2018; @bernard2021]. Dedicated seabird effort could thus benefit future studies advancing seabird research in this region, as well as improving patterns described here. By increasing the amount of data, future studies could set the grid cells at higher resolution, and potentially reveal assemblages at a finer scale. In addition, ideally, we could have had more information within each grid cell by season (our sampling units). Nonetheless, rarefaction curves show that the sample coverage [see @chao2012] was high at the RCP level, although there is room for detecting more species on most of them (**Fig S8**).
